```{r}
library(tidyverse)
library(ggplot2)
```

```{r}
wnba <- read_csv('wnba.csv')
```

### Proportions with Bar chart

```{r}
wnba%>%
ggplot(aes(x = Exp_ordinal, y = ..prop.. *100,
       group = 1, fill = factor(..x..)))+
geom_bar()
theme(legend.position = 'none')+
labs(x = 'Experience Level',
     y = 'Percentage')
```

### stacked bar chart
```{r}
pos_prop <- wnba %>% 
  group_by(Pos) %>% 
  summarize(Prop = n() / nrow(wnba))

ggplot(data = pos_prop, 
              aes(x = "", y = Prop, fill = Pos)) + 
  geom_bar(stat = "identity", width = 0.25) +
  coord_flip() +
  geom_text(aes(label = str_c(round(Prop * 100), "%")), 
            position = position_stack(vjust = 0.5)) + 
  labs(x = NULL, 
       y = NULL, 
       fill = NULL, 
       title = "Player Distribution by Position") + 
  theme_classic() + 
  theme(axis.line = element_blank(),
          axis.text = element_blank(),
          axis.ticks = element_blank())
```
### Pie chart

```{r}
pos_prop <- wnba %>% 
  group_by(Pos) %>% 
  summarize(Prop = n() / nrow(wnba))

ggplot(data = pos_prop, 
              aes(x = "", 
                  y = Prop, 
                  fill = Pos)) + 
  geom_bar(stat = "identity") +
  coord_polar(theta = "y")
```
### Histogram

A histogram is basically the visual form of a grouped frequency table. Structurally, a histogram can also be understood as a modified version of a bar chart. The main difference is that, in the case of a histogram, there are no gaps between bars, and each bar represents an interval, not a single value.

```{r}
games_binwidth <- (max(wnba$Games_Played) - min(wnba$Games_Played)) / 10
wnba%>%
ggplot(aes(x = Games_Played))+
geom_histogram(binwidth = games_binwidth,
               boundary = min(wnba$Games_Played))+
scale_x_continuous(breaks = seq(0,35,by=5))
```

### Comparing frequency distributions

```{r}

library(ggplot2)
wnba%>%
ggplot(aes(x = Pos, fill = Exp_ordinal))+
geom_bar(position = 'dodge')+
labs(x='Position')


pos_by_exp <- wnba %>% 
  group_by(Exp_ordinal, Pos)%>%
  summarise(Freq = n())

pos_by_exp%>%
ggplot(aes(x = Pos, y = Freq ,fill = Exp_ordinal))+
geom_bar(position = 'dodge',
        stat = 'identity')+
labs(x='Position')

```

### Challenge: Do Older Players Play Less?


```{r}
wnba <- wnba %>% 
  mutate(age_relative = 
         if_else(Age >= 27, "old", "young")) %>% 
  mutate(age_relative = 
         factor(age_relative, 
                levels = c("young", "old")))

wnba <- wnba %>% 
  mutate(min_relative = 
         if_else(MIN >= 497, "average or above", "below average")) %>% 
  mutate(min_relative = 
         factor(min_relative,
                levels = c("below average", "average or above")))

wnba %>%
ggplot( aes(x=age_relative, fill = min_relative))+
geom_bar(position = 'dodge')
```

The information provided by the MIN variable is much more granular, and we can plot the distribution of this variable instead. Because the MIN variable is measured on a ratio scale, we'll need to use histograms instead of bar plots. 


```{r}
ggplot(data = wnba,
         aes(x = MIN, fill = age_relative)) +
  geom_histogram(bins = 10, 
                 position = "identity", # Superimpose one on top of the other.
                 alpha = 0.5)+ # Make each histogram transparent
    geom_vline(aes(xintercept = mean(MIN), 
                   linetype = "Average minutes"), 
               color = "black")
```
```{r}
ggplot(data = wnba,
         aes(x = MIN, fill = age_relative)) +
  geom_histogram(bins = 10, 
                 position = "identity", # Superimpose one on top of the other.
                 alpha = 0.5)+ # Make each histogram transparent
    geom_vline(aes(xintercept = mean(MIN), 
                   linetype = "Average minutes"), 
               color = "black") +
  facet_wrap(~age_relative)
```

### Frequency Polygons

```{r}
ggplot(data = wnba,
         aes(x = MIN, color = age_relative)) +
  geom_freqpoly(bins = 10, 
                 position = "identity") +
  geom_vline(aes(xintercept = mean(MIN), 
                 linetype = "Average minutes"), 
             color = "black")
```

### Kernel Density Estimate Plots

```{r}
wnba%>%
ggplot(aes(x=MIN, color = age_relative))+
geom_density()+
geom_vline(aes(xintercept = mean(MIN),
               linetype = 'Average Minutes'),
           color='black')

```

### Drawbacks of Kernel Density Plots 

If we look very closely, we can see a couple of clear patterns: the shortest players are generally guards, the tallest players are generally centers, mid-height players are generally forwards or play in a combined position, etc.

Having to look very closely to a graph to identify obvious patterns is far from ideal. If there's any pattern, we want to see it immediately. 

```{r}
ggplot(data = wnba,
         aes(x = Height, color = Pos)) +
  geom_density()
```

### Scatter Plots (Strip Plots)

One alternative we can use to visualize the distribution of heights as a function of player position is this style of scatterplot where each category is represented independently on the x-axis:

```{r}
ggplot(data = wnba,
         aes(x = Pos, y = Height, color = Pos)) + 
  geom_point()
```

A downside of this scatter plot is that the bullet points overlap. We can fix this by adding a bit of jitter to each distribution. Jitter spreads the points out by applying a small amount of random noise to each point. We can do this by adding the geom_jitter() layer to the code:

```{r}
ggplot(data = wnba,
         aes(x = Pos, y = Height, color = Pos)) + 
  geom_point()+
  geom_jitter()
```

### Box Plots

Besides scatter plots, there's another kind of graph we can use to display many distributions at once and make sure everything is still readable.

```{r}
wnba%>%
ggplot(aes(x = Pos, y =Weight, color= Pos))+
geom_boxplot()
```
The few dots we see for the box plots of centers and guards/forwards (G/F) represent values in the distribution that are much larger or much lower than the rest of the values. A value that is much lower or much larger than the rest of the values in a distribution is called an outlier.

A value is an outlier if:

* It's larger than the upper quartile by 1.5 times the difference between the upper quartile and the lower quartile (the difference is also called the interquartile range).
* It's lower than the lower quartile by 1.5 times the difference between the upper quartile and the lower quartile (the difference is also called the interquartile range).

```{r}
sum(wnba$Games_Played > 0) # true value will count as 1 in the summation
sum(wnba$Games_Played)
nrow(wnba)
```
```{r}
compute_mode <- function(vector) {
    counts_df  <-  tibble(vector) %>% 
        group_by(vector) %>% 
        summarise(frequency = n()) %>% 
        arrange(desc(frequency)) 

    counts_df$vector[1]
}
ggplot(data = wnba,
    aes(x = Age)) +
    geom_density(alpha = 0.1, 
                 color='blue', 
                 fill='blue') +
    geom_vline(aes(xintercept = compute_mode(Age), 
                   color = 'Mode'), 
               size = 1.2 ) +
    geom_vline(aes(xintercept = median(Age), 
                   color = 'Median'), 
               size = 1.2 ) +
    geom_vline(aes(xintercept = mean(Age), 
                   color = 'Mean'), 
               size = 1.2 ) +
    scale_y_continuous(labels = scales::comma) +
    xlim(20,40)+
    scale_colour_manual(values = c("Mode" = "green", 
                                   "Median" = "black", 
                                   "Mean" = "orange"), 
                        name = "") +
    theme_bw() + 
    theme(legend.position='top') +
    xlab("Age") + 
    ylab("Density")
```
```{r}

```
